\begin{abstract}
Nowadays an electric grid is an important part of our society today. The electric grid consist of power plants, power lines and substations. With these building blocks electricity can be delivered from power plants to end users power plugs in the house. Substations and their automation play an important role of power grid safety and making sure it is working. The focus of this master thesis is to plan and implement a software component to be part of the bigger system related to substations and their management. Planned component should be able to subscribe for information from the substation and share that with other parts of the system. Information coming from the substation can for example include measurement data. Part of the system interested in this data should have an access to this data and then for example show it on user interface.

In substations the information is subscribed from an \emph{Intelligent Electronic Device}, IED for short. an IED is a automation device which controls the other physical devices of the substation. IEDs are also connected to the substation local network. IED can also be called with name of protection relay. The \emph{International Electrotechnical Commission} has defined a worldwide standard called \emph{IEC 61850}, which defines the rules how IED devices should communicate with each other over the substation network. This standard also defines rules how a software outside of the substation network can subscribe information from the IED. This part of the standard is used by the implemented software component to subscribe for the information.

Before star of this thesis, the creator had done a proof of concept software component. This component however had many problems in it and was a reason why it could not be used as part of the other system in production. Analyzing the problems was part of this thesis and the knowledge gained from this was used to make decisions regarding planning of the new software component.

As a result from this thesis was a software component independent from the other system. Implemented component was able to subscribe for the information from the IED according to IEC 61850 standard and share it with other parts of the system. Component ended up to the production environment and being part of the bigger system.
\end{abstract}